---
title: "Assignment 6, lab 3, group 6"  
author: Juan Moreno Diez, Letícia Marçal Russo, Luc Lubbers   
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
       highlight: textmate
       theme: flatly
       number_sections: yes
       toc: yes
       toc_float:
         collapsed: yes
         smooth_scroll: no
---
# Libraries

```{r libraries, warning = FALSE, message = FALSE}
library(tidyverse)
library(factoextra)
library(GGally)
library(ggplot2)
library(patchwork)
```

# Data

Here we will load the data. 

```{r data, warning = FALSE, message = FALSE}
clus_df <- read_csv("https://raw.githubusercontent.com/vankesteren/dav_practicals/master/12_Unsupervised_learning_Clustering/data/clusterdata.csv")
```

Checking the dataset bellow.

```{r head}
head(clus_df) %>% 
  knitr::kable(format = "markdown", digits= 1, padding = 30, align = 'c')
```

# Function for Euclidean distance

We will create a function to calculate the Euclidean distance. 

```{r}
l2_list <- function(x, y) {
  sqrt(sum((x - y)^2))
}
```

# K-medians clustering algorithm

Here we will program a K-medians clustering algorithm/ function called kmedians.

```{r}
kmedians <- function(dataframe, k) {
  # random inicialization of the centroids
  centroids <- dataframe[sample(nrow(dataframe), k), ]
  # default valur for number of iterations is 10
  n_iter <- 10
  
  # for each iteration 
  for (it in 1:n_iter) {
    # empty dataframe to fill with the distances
    distances <- data.frame(matrix(NA, nrow = nrow(dataframe), ncol = k))
    for (sample in 1:nrow(dataframe)) {
      for (centroid in 1:nrow(centroids)) {
        # euclidean distance between sample in dataframe and centroid
        distances[sample, centroid] <- l2_list(dataframe[sample, ], centroids[centroid, ])
      }
    }
    
    # assign the point to the minimun distance centroid
    assigned_clusters <- apply(distances, 1, which.min)
    
    # recalculate centroids
    for (i in seq_len(k)) {
      i_cluster <- dataframe[assigned_clusters == i,]
      # compute median
      centroids[k, ] <- as.list(apply(i_cluster, 2, median))
    }
  }
  return (list(centroids, assigned_clusters))
}
```

Running our algorithm with k = 7.
```{r}
set.seed(123)

list_centers <- kmedians(clus_df, k = 7)

# store in the dataframe the clusters assigned
# to each point by the k medians algorithm
clus_df['cluster_kmedians'] <- list_centers[2]

```

Information about the centroids. 
```{r}
print("Centroids found")
print(list_centers[1])

print("Clusters assigned to each point in the dataframe")
print(list_centers[2])

``` 

# Comparison to K-means

Now we will compare the results of K-medians against K-means. 
```{r}
# store in the dataframe the clusters assigned
# to each point by the k means algorithm
km.res <- kmeans(clus_df, 7, nstart = 100)
clus_df['cluster_kmeans'] <- km.res$cluster
```

We added a new column to the data frame with the clusters assigned by k-means. Bellow is the head of the dataset. 
```{r}
head(clus_df) %>% 
  knitr::kable(format = "markdown", digits= 1, padding = 30, align = 'c')
```

```{r plots, warning = FALSE, message = FALSE}
median_plot <- clus_df %>%
  ggplot() +
    geom_point(aes(x = x1, y = x2, 
                   colour = as.factor(cluster_kmedians))) +
    scale_fill_brewer(palette = "Dark2") +
    guides(col = FALSE) +
    ggtitle("K-median") +
    theme_minimal() 

mean_plot <- clus_df %>%
  ggplot() +
    geom_point(aes(x = x1, y = x2, 
                   colour = as.factor(cluster_kmeans))) +
    scale_fill_brewer(palette = "Dark2") +
    guides(col = FALSE) +
    ggtitle("K-means") +
    theme_minimal() 

median_plot + mean_plot

```


The plots above shows that the main difference in clusters between median and mean is in the top of the plot (x2 above 0.5). The cluster on the bottom it is pretty much the same in both plots, probably because the points keep certain distance in relation to the top points. 

On the other hand, on the top of the plots, the points are more concetrated and then one can see the difference between median and mean more clearly. F




